---
title: "reading_data_from_the_web"
output: github_document
date: "2025-10-20"
---

```{r setup, include=FALSE}

library(tidyverse)
library(rvest)
library(httr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

Import NSDUH data from the web

```{r}
url = "https://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)
```
Defining the url and then reading in the url as html.


This is an "easy" case

```{r}
ndsuh_df =
drug_use_html |> 
  html_table() |> 
  first() |> 
  slice(-1)
```

`first()` gives you the first table (the first thing in the list)
you can also use `nth()` and insert the table number in ()
`slice()` can get rid of the row 


Slightly harder case

```{r}
url = "https://www.imdb.com/list/ls070150896/"

sw_html =
  read_html(url)
```

Now pull out elements of the html that I care about using SelectorGadget

```{r}
title_vec =
sw_html |> 
  html_elements(".ipc-title-link-wrapper .ipc-title__text--reduced") |> 
  html_text()

metascore_vec =
  sw_html |> 
  html_elements(".metacritic-score-box") |> 
  html_text()

runtime_vec =
  sw_html |> 
  html_elements(".dli-title-metadata-item:nth-child(2)") |> 
  html_text()

sw_df =
  tibble(
    title = title_vec,
    metascore = metascore_vec,
    runtime = runtime_vec
  )
```


## APIs

Get data using an API
```{r}
nyc_water_df =
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") |> 
  content("parsed")

nyc_water_df |> 
  ggplot(aes(x = year, y = nyc_consumption_million_gallons_per_day))+
  geom_point()
```
`content(parsed)` make a guess about file structure and what it looks like


Access BRFSS

```{r}
brfss_df =
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv") |> 
  content("parsed")
```
If you need to access your data and it requires a token you will add inside GET():
query = list("app_token"= "insert token code")


Look at Pokemon data
```{r}
poke =
  GET("http://pokeapi.co/api/v2/pokemon/ditto") |>
  content()

poke[[4]]
```


