---
title: "Strings and Factors"
output: github_document
date: "2025-10-21"
---

Load key packages.
```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
```

## Mostly use string vectors

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
str_detect(string_vec, "e")
str_detect(string_vec, "Jeff")

str_replace(string_vec, "jeff", "Jeff")
str_replace(string_vec, "e", "E")
str_replace(string_vec, "e", "")
```
`str_detect` goes into find the string that you are looking for. Case sensitivity also matters.



copied from p8105.com

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")

str_remove(string_vec, "i think$")
```
Using "^" indicates start of the string
Using "$" indicates end of the string
`str_remove` removes part of string u want.


Pumpkin themed example!
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "Pumpkin")
str_detect(string_vec, "[Pp]umpkin")
```
[ ] indicates anything you want to match inside strings. 



Let's get a bit more complicated. (Pattern Matching Problem)
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

[0-9]: anything between 0 and 9 counts
[A-Z]: anything between A and Z counts



Another example
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.1")
str_detect(string_vec, "7\\.1")
```

7.1: the "." means anything that matches to it. 
If you are trying to include a special character like "." in your detection, insert "\\"



## Factors

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex

as.numeric(vec_sex)

vec_sex = fct_inorder(vec_sex)
vec_sex
```

## Revisit NSDUH

Import NSDUH data from the web

```{r}
url = "https://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)
```
Defining the url and then reading in the url as html.


This is an "easy" case

```{r}
nsduh_df =
drug_use_html |> 
  html_table() |> 
  first() |> 
  slice(-1)
```
`first()` gives you the first table (the first thing in the list)
you can also use `nth()` and insert the table number in ()
`slice()` can get rid of the row 


Now we do the tidying...
```{r}
marj_df =
  nsduh_df |> 
  select(-contains("P Value")) |> 
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent"
  ) |> 
  separate(age_year, into = c("age", "year"), sep = "\\(") |> 
  mutate(
    year = str_remove(year, "\\)"),
    percent = str_remove(percent, "[a-c]$"),
    percent = as.numeric(percent)
  ) |> 
  filter(
    !State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")
  )
```


Let's make a quick plot
```{r}
marj_df |> 
  filter(age == "12-17") |> 
  mutate(State = fct_reorder(State, percent)) |> 
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



