---
title: "Strings and Factors"
output: github_document
date: "2025-10-21"
---

Load key packages.
```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
```

## Mostly use string vectors

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
str_detect(string_vec, "e")
str_detect(string_vec, "Jeff")

str_replace(string_vec, "jeff", "Jeff")
str_replace(string_vec, "e", "E")
str_replace(string_vec, "e", "")
```
`str_detect` goes into find the string that you are looking for. Case sensitivity also matters.



copied from p8105.com

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")

str_remove(string_vec, "i think$")
```
Using "^" indicates start of the string
Using "$" indicates end of the string
`str_remove` removes part of string u want.


Pumpkin themed example!
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "Pumpkin")
str_detect(string_vec, "[Pp]umpkin")
```
[ ] indicates anything you want to match inside strings. 



Let's get a bit more complicated. (Pattern Matching Problem)
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

[0-9]: anything between 0 and 9 counts
[A-Z]: anything between A and Z counts



Another example
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.1")
str_detect(string_vec, "7\\.1")
```

7.1: the "." means anything that matches to it. 
If you are trying to include a special character like "." in your detection, insert "\\"



## Factors

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex

as.numeric(vec_sex)

vec_sex = fct_inorder(vec_sex)
vec_sex
```










